---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Packages
```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(chron)
```


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity <- read_csv("activity.csv", col_types = cols(date = col_date(format = "%Y-%m-%d"), 
    steps = col_integer()))
```



## What is mean total number of steps taken per day?

1. Calculate Total Number of steps taken per day
```{r}
total<-aggregate(steps~date,activity, sum)
```

2.Histogram of total number of steps taken each day
```{r}
ggplot(total, aes(x=steps)) +geom_histogram(bins=5)+ labs(title="Total Steps Per Day", x="Steps", y="Frequency")
```

3. Mean and Median number of steps taken per day  
-Mean
```{r}
mean(total$steps)
```
-Median
```{r}
median(total$steps)
```


## What is the average daily activity pattern?
1. Calculate Average number of steps taken per interval
```{r}
avgperint<-aggregate(steps~interval, activity, mean)
```
Create Time series plot
```{r}
plot(avgperint$interval, avgperint$steps,  type="l", main="Avg steps per interval", xlab="interval", ylab="steps")
```  

2. Interval with maximum number of steps
Arrange the intervals from greatest number of steps to least
```{r}
max<-arrange(avgperint, desc(avgperint$steps))
```
The interval with the maximum number of steps
```{r}
max[1,1]
```



## Imputing missing values  
1. Total number of missing values
```{r}
sum(is.na(activity$steps))
```  
2. Fill in the missing values using the mean value for the interval
```{r}
merged<-merge(activity, avgperint, by="interval")
merged$stepsimputed=ifelse(is.na(merged$steps.x), merged$steps.y, merged$steps.x)
```  
3. Create new data set
```{r}
imputed<-merged[, c(1, 3, 5)]
```  
4. Create histogram of total steps per day  
-Calculate total steps per day
```{r}
imputedtotal<-aggregate(stepsimputed~date,imputed, sum)
```  
-create histogram
```{r}
ggplot(imputedtotal, aes(x=stepsimputed)) +geom_histogram(bins=5)+ labs(title="Total Steps Per Day",subtitle="Uses average number of steps for interval for missing values", x="Steps", y="Frequency")
```    

-Mean number of steps per day
```{r}
mean(imputedtotal$steps)
```  
-Median number of steps per day
```{r}
median(imputedtotal$steps)
```  
The mean is the same whether we impute missing values or not.  
The median is slightly higher.  

## Are there differences in activity patterns between weekdays and weekends?  
1. Create factor variable for weekday or weekend.  
```{r}
first<-mutate(imputedtotal, weekend=is.weekend(imputedtotal$date)) 
second<-mutate(first, type=factor(weekend, levels=c("TRUE", "FALSE"), labels=c("weekend", "weekday")))
```


